---
title: "Technical Appendix: Age-Matched Demand Denominators"
subtitle: "Supplemental Digital Content"
output:
  word_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressWarnings(suppressMessages({library(knitr); library(yaml)}))
root <- tryCatch(rprojroot::find_root(rprojroot::has_file("DESCRIPTION")), error = function(e) "..")
knitr::opts_knit$set(root.dir = root)
fmt  <- function(x, d = 0) formatC(x, format = "f", digits = d, big.mark = ",")
rat  <- function(x) formatC(x, format = "f", digits = 4)
```

```{r load}
MAN  <- file.path(root, "inst/multiverse/age_matched_denominator.yml")
HASH <- file.path(root, "inst/multiverse/age_matched_denominator.sha256")
man  <- yaml::read_yaml(MAN)
man_sha <- trimws(strsplit(readLines(HASH, warn = FALSE)[1], "\\s+")[[1]][1])

amd_path <- file.path(root, "artifacts/2sfca/sensitivity/cache/age_matched_denominators.rds")
amd <- if (file.exists(amd_path)) readRDS(amd_path) else NULL

# The PANEL is authoritative: eleven vintages computed end to end on the frozen
# AMI. Its 2020 rows are byte-identical to the standalone file, but reading the
# panel means this appendix, the manuscript and the correction diff all cite one
# artifact rather than three that happen to agree.
# MANDATORY. No fallback to the standalone 2020 file -- see the manuscript setup
# chunk for why fail-open provenance is worse than a hard stop.
res_path <- file.path(root, "artifacts/2sfca/agematched_panel/age_matched_panel.csv")
if (!file.exists(res_path))
  stop("the age-matched panel is missing: ", res_path,
       "\n  It is the sole source for every age-matched number in this appendix.")
res <- utils::read.csv(res_path, stringsAsFactors = FALSE)
if (!is.null(res) && "year" %in% names(res)) res <- res[res$year == 2020L, , drop = FALSE]
have_results <- !is.null(res) && nrow(res) > 0
```

### 1. The problem this appendix addresses

In the primary analysis every subspecialty is indexed to the same demand
denominator: total female population of all ages (ACS 2020, table `B01001`,
variable `B01001_026`; 164,690,617 women across the contiguous United States).

That choice is defensible as a *standardized* measure of geographic
reachability, and the manuscript says so. But it has a consequence that a
careful reader will notice immediately:

- **pediatric and adolescent gynecology** is measured against a population that
  includes 80-year-old women;
- **gynecologic oncology** and **urogynecology** are measured against a
  population that includes 6-year-old girls.

Because accessibility is supply divided by weighted demand, a denominator
containing large numbers of people who are not candidates for a given
subspecialty deflates that subspecialty's apparent access. The finding that
complex family planning and pediatric and adolescent gynecology have the lowest
access is therefore, in part, an artifact of who is counted in the denominator
rather than a statement about service availability.

This appendix reports a prespecified sensitivity analysis in which each
subspecialty is indexed to an age range plausibly served by that subspecialty.

### 2. Prespecification

The age windows, the ACS variables implementing them, the estimands, and the
interpretation rules were written to
`inst/multiverse/age_matched_denominator.yml` and recorded by SHA-256
(`inst/multiverse/age_matched_denominator.sha256`) **before any outcome was
computed**. The analysis code verifies that hash before reading data and aborts
if the manifest has changed. Manifest versions were incremented rather than
edited in place; the history of revisions, and the reason for each, is in the
file itself.

Age ranges are a clinical judgement, contributed by the senior author, and are
recorded as such rather than derived from the data.

### 3. Age windows

| Subspecialty | Age window | Clinical intent | ACS bands (total) |
|---|---|---|---|
| Pediatric and adolescent gynecology | under 20 | under 21 | `B01001_027`–`_031` |
| Complex family planning | under 35 | under 40 | `B01001_027`–`_036` |
| Gynecologic oncology | 45 and over | — | `B01001_039`–`_049` |
| Urogynecology and reconstructive pelvic surgery | 45 and over | — | `B01001_039`–`_049` |
| Maternal-fetal medicine | 15 to 44 | — | `B01001_030`–`_038` |
| Reproductive endocrinology and infertility | 15 to 44 | — | `B01001_030`–`_038` |
| Minimally invasive gynecologic surgery | 15 and over | — | `B01001_030`–`_049` |

### 4. Why two windows differ from the clinical intent

The disparity contrasts re-weight the same accessibility surface by subgroup
population. Rurality is assigned at tract level (RUCA), so the rural and
metropolitan layers can use the fine-grained `B01001` table, which has 23 female
age bands. The race-specific layers must use the ACS race iterations
`B01001H` (White alone, not Hispanic) and `B01001C` (American Indian and Alaska
Native alone), which have only **14** female age bands and are decadal above age
35.

The band edges common to all three tables are:

```
0  5  10  15  18  20  25  30  35  45  55  65  75  85
```

Ages 40, 50, 60, 70 and 80 do not exist as boundaries in the race iterations.
Five of the seven clinically-specified windows already fell on common edges. Two
did not:

- **pediatric and adolescent gynecology**, intended as under 21, was moved to
  **under 20** (the adjacent common edge, one year away; the alternative was 25);
- **complex family planning**, intended as under 40, was moved to **under 35**
  (the two adjacent common edges, 35 and 45, are equidistant).

The alternative to shifting was to interpolate within a straddling band, which
would impose a uniform-age-distribution assumption precisely where the American
Indian and Alaska Native cell counts are smallest and least stable. Shifting to
a common edge means every contrast in every subspecialty uses one age window,
so the rurality and race comparisons are directly comparable rather than
requiring reconciliation.

**Direction of the complex family planning choice.** Under-35 is the narrower of
the two equidistant options. A narrower demand denominator raises apparent
access. Complex family planning reports the lowest access of the seven
subspecialties in the primary analysis, so this choice moves that estimate in
the direction that makes it look better. This was recorded in the manifest
before results were computed.

### 4a. Formal statement: where the denominator enters, and what it can change

Let $i$ index population cells on the analysis grid and $j$ index provider
locations. Write $P_i$ for the demand population in cell $i$, $S_j$ for the
supply at provider $j$, and $w_{ij}$ for the incremental distance-decay weight
linking them. The two steps of E2SFCA are

$$R_j \;=\; \frac{S_j}{\sum_i w_{ij} P_i}, \qquad\qquad A_i \;=\; \sum_j w_{ij} R_j .$$

The demand denominator enters **only** through $\sum_i w_{ij} P_i$ in step 1.
Age-matching replaces $P_i$ with

$$P_i^{(s)} \;=\; \alpha_i^{(s)} \, P_i ,$$

where $\alpha_i^{(s)} \in [0,1]$ is the share of cell $i$'s female population
falling inside subspecialty $s$'s age window. Nationally,
$\alpha^{(s)} = \sum_i P_i^{(s)} / \sum_i P_i$ is the quantity plotted in
Figure S1, Panel A: 0.24 for pediatric and adolescent gynecology, 0.82 for
minimally invasive gynecologic surgery.

The reported subgroup mean for population group $G$ (rural, metropolitan,
White, American Indian and Alaska Native) is the population-weighted average

$$\bar{A}_G \;=\; \frac{\sum_i P_i^{G} A_i}{\sum_i P_i^{G}},$$

and each reported contrast is a ratio of two such means, for example
$\bar{A}_{\text{rural}} / \bar{A}_{\text{metro}}$.

#### Why levels are not comparable, and ratios are

**Proposition.** If $\alpha_i^{(s)} = \alpha$ for every cell $i$ — that is, if
the age-eligible share were spatially uniform — then

$$R_j^{(s)} = \frac{S_j}{\sum_i w_{ij}\,\alpha P_i} = \frac{1}{\alpha} R_j,
\qquad A_i^{(s)} = \frac{1}{\alpha} A_i,
\qquad \bar{A}_G^{(s)} = \frac{1}{\alpha} \bar{A}_G ,$$

so every accessibility level is inflated by the constant factor $1/\alpha$ while
**every contrast ratio is exactly unchanged**, because the factor cancels:

$$\frac{\bar{A}_{G_1}^{(s)}}{\bar{A}_{G_2}^{(s)}}
= \frac{\alpha^{-1}\bar{A}_{G_1}}{\alpha^{-1}\bar{A}_{G_2}}
= \frac{\bar{A}_{G_1}}{\bar{A}_{G_2}} .$$

```{r invariance-check, results="asis"}
if (have_results && !is.null(amd)) {
  k  <- "GO"
  n0 <- res$national[res$regime == "all_ages"    & res$subspec == k]
  n1 <- res$national[res$regime == "age_matched" & res$subspec == k]
  W1 <- readRDS(file.path(root, "artifacts/2sfca/sensitivity/cache/acs2020_age_bands.rds"))
  al <- sum(amd$denominators[[k]]$total_f) / sum(W1$B01001_026)
  cat(sprintf(paste0("This is the formal content of the interpretation rule in section 7. ",
    "Gynecologic oncology's national mean moves from %s to %s, an observed factor of ",
    "**%s**, against a predicted $1/\\alpha = 1/%s = %s$. It is arithmetic, not a finding.\n"),
    rat(n0), rat(n1), rat(n1 / n0), rat(al), rat(1 / al)))
}
```

**Corollary — what the analysis actually tests.** Because the uniform case
leaves contrasts exactly unchanged, *any* movement in a contrast under
age-matching is attributable entirely to **spatial and between-group variation
in $\alpha_i$**. The analysis is therefore not a test of whether access differs,
but of whether the age composition of the population covaries with the geography
of access, or with group membership, strongly enough to move a disparity.

#### A testable prediction, and its result

The two race groups have materially different age structures nationally:

```{r age-structure, results="asis"}
wp0 <- file.path(root, "artifacts/2sfca/sensitivity/cache/acs2020_age_bands.rds")
if (!is.null(amd) && file.exists(wp0)) {
  W0 <- readRDS(wp0)
  dw <- sum(W0$B01001H_017); da <- sum(W0$B01001C_017)
  wins <- c(PAG = "under 20", GO = "45 and over")
  out <- do.call(rbind, lapply(names(wins), function(k) {
    d  <- amd$denominators[[k]]
    aw <- sum(d$white_f) / dw; aa <- sum(d$aian_f) / da
    data.frame(`Age window` = unname(wins[k]),
               `White NH share` = rat(aw), `AIAN share` = rat(aa),
               `rho = alpha_AIAN / alpha_White` = rat(aa / aw), check.names = FALSE) }))
  print(knitr::kable(out, align = "lrrr"))
}
```

The American Indian and Alaska Native female population is younger. Age
restriction therefore does not scale the two groups equally, and the direction
of the effect is predictable. When $\rho > 1$ the age window retains
proportionally *more* AIAN demand, which raises the demand denominator in cells
where those women live, lowering $R_j$ and hence their measured access relative
to White-area access. When $\rho < 1$ the reverse holds. Formally, the contrast
should move in the direction of $-\operatorname{sign}(\rho - 1)$.

```{r rho, results="asis"}
amdp <- file.path(root, "artifacts/2sfca/sensitivity/cache/age_matched_denominators.rds")
wp   <- file.path(root, "artifacts/2sfca/sensitivity/cache/acs2020_age_bands.rds")
if (have_results && file.exists(amdp) && file.exists(wp)) {
  A <- readRDS(amdp); W <- readRDS(wp)
  wht <- sum(W$B01001H_017); ai <- sum(W$B01001C_017)
  aa2 <- res[res$regime=="all_ages",];  rownames(aa2) <- aa2$subspec
  am2 <- res[res$regime=="age_matched",]; rownames(am2) <- am2$subspec
  ord2 <- c("PAG","CFP","MFM","REI","MIGS","GO","FPMRS")
  tab <- do.call(rbind, lapply(ord2, function(k) {
    d <- A$denominators[[k]]
    rho <- (sum(d$aian_f)/ai) / (sum(d$white_f)/wht)
    obs <- am2[k,"aian_white_ratio"] - aa2[k,"aian_white_ratio"]
    data.frame(Subspecialty = k, rho = rat(rho),
               `AIAN:White, all ages` = rat(aa2[k,"aian_white_ratio"]),
               `AIAN:White, age-matched` = rat(am2[k,"aian_white_ratio"]),
               Predicted = ifelse(rho > 1, "widen", "narrow"),
               Observed  = ifelse(obs < 0, "widen", "narrow"),
               Agreement = ifelse((rho > 1) == (obs < 0), "yes", "no"),
               check.names = FALSE) }))
  print(knitr::kable(tab, align = "lrrrlll"))
  cat(sprintf("\n\nThe predicted direction is observed in **%d of %d** subspecialties.\n",
              sum(tab$Agreement == "yes"), nrow(tab)))
}
```

The changes in the American Indian and Alaska Native contrast are therefore not
noise: they are the mechanical consequence of a younger age structure
interacting with a narrower age window. In the 2020 sensitivity analysis the
disparity persists in every subspecialty, but age-matching *widens* it wherever
the window favours younger women and *narrows* it wherever the window favours
older women.

```{r cfp-parity, include=FALSE}
# Computed, never typed: this paragraph reports where the AIAN:White contrast
# crosses parity, and a typed value here would go stale the moment the panel is
# recomputed. The appendix literal guard forbids >= 3-decimal literals in prose
# for exactly this reason.
.pan <- utils::read.csv("artifacts/2sfca/agematched_panel/age_matched_panel.csv",
                        stringsAsFactors = FALSE)
.cfp <- .pan[.pan$subspec == "CFP", ]
.over <- .cfp[.cfp$aian_white_ratio >= 1, ]
.n_over_specs <- length(unique(.pan$subspec[.pan$aian_white_ratio >= 1]))
.yrs_over <- sort(unique(.over$year))
.lo <- .cfp[which.min(.cfp$aian_white_ratio), ]
.f2 <- function(x) sprintf("%.2f", x)
.get <- function(y, rg) .cfp$aian_white_ratio[.cfp$year == y & .cfp$regime == rg]
```

The 2020 scope of that statement is deliberate. Across the full eleven-year panel
the American Indian and Alaska Native contrast is **not** below parity in every
subspecialty-year. Complex family planning is the exception, and the only one:
it reaches its minimum of `r .f2(.lo$aian_white_ratio)` in `r .lo$year`, is still
below parity in 2020 at `r .f2(.get(2020, "age_matched"))` age-matched, and then
crosses parity in `r paste(.yrs_over, collapse = " and ")` --
`r .f2(.get(2022, "all_ages"))` and `r .f2(.get(2023, "all_ages"))` all-ages,
`r .f2(.get(2022, "age_matched"))` and `r .f2(.get(2023, "age_matched"))`
age-matched. The path there is not monotonic; it falls through
`r .lo$year` before rising. That `r .n_over_specs` of seven subspecialties ever
exceed parity is consistent with the specification curve, which already
identifies this same contrast as the fragile one -- it reverses for complex
family planning under flatter decay and under M2SFCA. Reporting the crossing
rather than suppressing it is the point: a disparity claim that held everywhere
and always would be a claim this analysis does not support.

#### Conservation

The engine satisfies the identity

$$\sum_i P_i A_i \;=\; \sum_j S_j$$

over providers reachable by some population — accessibility redistributes supply
across space without creating or destroying it. This holds separately under each
denominator regime, with $P_i^{(s)}$ substituted throughout; it is verified
numerically in the package test suite rather than assumed here.

### 4b. Supply conservation

Age-restricted accessibility estimates are only meaningful if the supply side is
complete. Every provider origin entering the model must have an associated
drive-time catchment. The pipeline fails closed if any does not, naming the
offending origins and quantifying the supply they carry, rather than dropping
them and continuing.

This is enforced rather than assumed because the failure mode is silent: supply
with no catchment contributes to no one's accessibility, so it deflates every
downstream mean while each individual value still looks plausible. An earlier
version of this analysis dropped unmatched origins by configuration and landed
below the frozen reference, with the loss visible only in a column that nothing
compared.

```{r supply-loss, results="asis"}
.sp <- file.path(root, "artifacts/2sfca/sensitivity/sensitivity_2020.csv")
# The CONTAMINATED artifact, deliberately. This paragraph describes the historical
# defect, and computing it from the corrected file made it report a 0.000%
# shortfall from 0 lost origins -- the defect described using data in which it had
# been fixed.
.am <- file.path(root, "artifacts/multiverse/_precorrection/age_matched_results_CONTAMINATED_2020.csv")
if (file.exists(.sp) && file.exists(.am)) {
  .s <- utils::read.csv(.sp, stringsAsFactors = FALSE)
  .a <- utils::read.csv(.am, stringsAsFactors = FALSE)
  .ref <- .s$national[.s$variant == "base" & .s$subspec == "GO"]
  .got <- .a$national[.a$regime == "all_ages" & .a$subspec == "GO"]
  .lost <- .a$n_supply_origins[.a$subspec == "GO"][1] - .a$n_iso_origins[.a$subspec == "GO"][1]
  .tot  <- .a$n_supply_origins[.a$subspec == "GO"][1]
  cat(sprintf(paste0("For gynecologic oncology that shortfall was **%s%%** ",
    "relative to the frozen reference, from %d of %d provider origins having no ",
    "catchment. All results reported here come from runs in which supply is ",
    "fully conserved.\n"),
    formatC(100 * (.ref - .got) / .ref, format = "f", digits = 3), .lost, .tot))
}
```

### 5. Validation of the denominators

Three checks were run before any accessibility computation.

**Conservation.** Summing every female age band must reproduce the published
female total for each table. Exact agreement in all three:

```{r conservation}
raw <- file.path(root, "artifacts/2sfca/sensitivity/cache/acs2020_age_bands.rds")
if (file.exists(raw)) {
  w <- readRDS(raw)
  spec <- list(
    list("B01001 (total)",     "B01001_026",  sprintf("B01001_%03d", 27:49)),
    list("B01001H (White NH)", "B01001H_017", sprintf("B01001H_%03d", 18:31)),
    list("B01001C (AIAN)",     "B01001C_017", sprintf("B01001C_%03d", 18:31)))
  tab <- do.call(rbind, lapply(spec, function(s) {
    tot <- s[[2]]; parts <- intersect(s[[3]], names(w))
    a <- sum(w[[tot]], na.rm = TRUE)
    b <- sum(rowSums(w[, parts, drop = FALSE], na.rm = TRUE))
    data.frame(Table = s[[1]], Published = fmt(a), `Sum of bands` = fmt(b),
               `Relative difference` = formatC(abs(a-b)/max(a,1), format="e", digits=2),
               check.names = FALSE) }))
  knitr::kable(tab, align = "lrrr")
} else cat("_ACS age-band extract not present; conservation table omitted._")
```

**Overlap identity.** The MIGS window (15 and over) and the PAG window (under
20) together span all ages but both contain ages 15–19. Their sum minus the
published female total must therefore equal the female population aged 15–19:

```{r overlap, results="asis"}
if (!is.null(amd)) {
  migs <- sum(amd$denominators$MIGS$total_f, na.rm = TRUE)
  pag  <- sum(amd$denominators$PAG$total_f,  na.rm = TRUE)
  raw  <- file.path(root, "artifacts/2sfca/sensitivity/cache/acs2020_age_bands.rds")
  tot  <- if (file.exists(raw)) sum(readRDS(raw)[["B01001_026"]], na.rm = TRUE) else NA
  b1519 <- if (file.exists(raw)) {
    w <- readRDS(raw); sum(rowSums(w[, intersect(c("B01001_030","B01001_031"), names(w)),
                                     drop = FALSE], na.rm = TRUE)) } else NA
  cat(sprintf("```\n%s + %s - %s = %s\nACS female aged 15-19%s= %s\n```\n",
              fmt(migs), fmt(pag), fmt(tot), fmt(migs + pag - tot),
              strrep(" ", 6), fmt(b1519)))
}
```

This check uses only published totals and is independent of the band lists.

**Partition.** In every tract, metropolitan plus rural population must not
exceed the total. RUCA leaves some tracts in neither class, so equality is not
required, but excess would indicate a classification error.

### 6. Analysis

Both denominator regimes — all-ages and age-matched — were computed in a single
run using **identical** grid geometry, isochrones, provider supply, distance-decay
weights (Luo and Qi 2009 weight set 1: 1.00, 0.68, 0.22, 0.09), grid resolution
(1,000 m, EPSG:5070) and allocation method (mass-conserving, area-weighted).
Only the demand denominator differs. Any incompleteness in the reproduction
inputs therefore affects both regimes equally and cancels in the contrast.

Per-subspecialty denominators (contiguous United States, ACS 2020):

```{r denominators}
if (!is.null(amd)) {
  ord <- c("PAG","CFP","GO","FPMRS","MFM","REI","MIGS")
  win <- setNames(vapply(man$subspecialties, function(s) s$age_range, character(1)),
                  vapply(man$subspecialties, function(s) s$code, character(1)))
  tab <- do.call(rbind, lapply(ord, function(k) {
    d <- amd$denominators[[k]]
    data.frame(Subspecialty = k, Window = win[[k]],
               Total = fmt(sum(d$total_f, na.rm=TRUE)),
               Metropolitan = fmt(sum(d$metro_f, na.rm=TRUE)),
               Rural = fmt(sum(d$rural_f, na.rm=TRUE)),
               `White NH` = fmt(sum(d$white_f, na.rm=TRUE)),
               AIAN = fmt(sum(d$aian_f, na.rm=TRUE)), check.names = FALSE) }))
  knitr::kable(tab, align = "llrrrrr")
} else cat("_Denominators not yet built._")
```

### 7. Interpretation rule, fixed in advance

Age-matched and all-ages accessibility values are **not comparable in level**.
They have different denominators and are therefore different estimands. A change
in the national mean between regimes is arithmetic, not a finding: halving the
denominator approximately doubles the reported value.

Only two things may be compared across regimes:

1. the **direction** of each contrast (rural versus metropolitan; American
   Indian and Alaska Native area versus White area);
2. whether each qualitative claim in the manuscript remains true.

### 8. Results

```{r results, results="asis"}
if (!have_results) {
  cat("_The analysis run has not yet completed. This section is generated from",
      "`artifacts/2sfca/agematched_panel/age_matched_panel.csv`; re-knit once it exists._\n")
} else {
  ord <- c("GO","FPMRS","MFM","REI","MIGS","PAG","CFP")
  res$subspec <- factor(res$subspec, levels = ord)
  res <- res[order(res$subspec, res$regime), ]
  tab <- data.frame(
    Subspecialty = as.character(res$subspec),
    Regime = ifelse(res$regime == "all_ages", "all ages", res$age_range),
    Denominator = fmt(res$denominator),
    `National mean` = rat(res$national),
    `Rural:metro` = rat(res$rural_metro_ratio),
    `AIAN:White` = rat(res$aian_white_ratio),
    check.names = FALSE)
  print(knitr::kable(tab, align = "llrrrr"))
  cat("\n\n**Claim survival.**\n\n")
  am <- res[res$regime == "age_matched", ]
  c2 <- all(am$rural_metro_ratio < 1, na.rm = TRUE)
  c3 <- all(am$aian_white_ratio  < 1, na.rm = TRUE)
  c3_fail <- as.character(am$subspec[am$aian_white_ratio >= 1])
  aa <- res[res$regime == "all_ages", ]
  ord_aa <- as.character(aa$subspec[order(-aa$national)])
  ord_am <- as.character(am$subspec[order(-am$national)])
  c5 <- identical(ord_aa, ord_am)
  cat(sprintf("- **C2** (rural below metropolitan in all seven): **%s**\n",
              if (c2) "holds" else "FAILS"))
  cat(sprintf("- **C3** (AIAN area below White area in all seven): **%s**%s\n",
              if (c3) "holds" else "FAILS",
              if (length(c3_fail)) paste0(" — reverses for ", paste(c3_fail, collapse=", ")) else ""))
  cat(sprintf("- **C5** (subspecialty ordering stable): **%s**\n", if (c5) "holds" else "FAILS"))
  cat(sprintf("\n  - ordering, all ages:   %s\n", paste(ord_aa, collapse = " > ")))
  cat(sprintf("  - ordering, age-matched: %s\n", paste(ord_am, collapse = " > ")))
}
```

### Figure S1

```{r fig, out.width="100%", fig.cap="Age-matched versus all-ages demand denominators. Panel A: the share of the all-ages female population retained by each subspecialty's age window; all-ages indexing measures every subspecialty against 100 percent. Panel B: the two disparity contrasts, paired within subspecialty. Ratios are dimensionless and therefore comparable across regimes, unlike the underlying levels. Panel C: rank by national mean access; rank is comparable across regimes even though the values are not."}
# knitr and pandoc resolve this path from DIFFERENT directories, which is why
# every single-path attempt failed:
#   knitr  resolves from opts_knit$root.dir  -> the repo root
#   pandoc resolves from the Rmd's directory -> manuscript/
# So the existence check uses the repo-root path and the emitted path is
# Rmd-relative, with error = FALSE so include_graphics does not re-check it
# against the wrong directory.
#
# The figure lives beside the document rather than in artifacts/, so a submitted
# supplement carries its own figure.
check <- "manuscript/figures/fig_age_matched_denominators.jpg"   # for file.exists
emit  <- "figures/fig_age_matched_denominators.jpg"              # for pandoc
if (file.exists(check)) knitr::include_graphics(emit, error = FALSE) else
  knitr::asis_output("_Figure not yet generated._")
```

```{r fig-omission, results="asis"}
go0 <- if (have_results) res$national[res$regime == "all_ages"    & res$subspec == "GO"] else NA
go1 <- if (have_results) res$national[res$regime == "age_matched" & res$subspec == "GO"] else NA
cat(sprintf(paste0("**Note on what this figure deliberately omits.** National mean access is ",
  "not plotted for the two regimes side by side. Those values have different denominators ",
  "and are different estimands; gynecologic oncology moves from %s to %s under age-matching, ",
  "which is arithmetic rather than a change in access. A chart placing them together would ",
  "invite precisely the misreading section 7 forbids.\n"), rat(go0), rat(go1)))
```

**Anticipated in advance:** age-matching can legitimately reorder the
subspecialties, because the ordering in the primary analysis partly reflects how
much of the all-ages female population is a plausible candidate for each
subspecialty. A reordering is a result of this analysis, not a failure of it.

### 9. Limitations

**Age is a proxy for candidacy, not a definition of it.** A 50-year-old woman
may require complex family planning services; a 30-year-old may require
gynecologic oncology. These windows describe the population a subspecialty
predominantly serves, not an eligibility rule.

**The race contrasts are constrained by ACS granularity.** The race iterations
are decadal above age 35, which is why two windows were shifted. Finer
age-matching is available for the rurality contrast than for the race contrast;
using the finer resolution only where it exists would produce contrasts that are
not comparable to one another, and was deliberately not done.

**Area-level race measurement is unchanged.** As in the primary analysis, race
contrasts compare accessibility in areas where American Indian and Alaska Native
women live against areas where White women live. They are ecological measures
and do not describe individual experience.

**The American Indian and Alaska Native denominator remains small.** At
1,296,971 women nationally against 99,126,561 White women — a 76-fold
difference — these estimates are the least stable in the study, before and after
age-matching.

**Potential, not realized, access.** Age-matching addresses who is counted in
the denominator. It does not address insurance acceptance, appointment
availability, transportation, or whether care is sought.

### 10. Reproducibility

| Component | Path |
|---|---|
| Frozen specification | `inst/multiverse/age_matched_denominator.yml` |
| Specification hash | `inst/multiverse/age_matched_denominator.sha256` |
| Denominator construction | `tools/multiverse/age_matched_denominators.R` |
| Analysis | `tools/multiverse/run_age_matched.R` |
| Results | `artifacts/2sfca/agematched_panel/age_matched_panel.csv` (11 vintages, frozen AMI) |
| Historical defect example | `artifacts/multiverse/_precorrection/age_matched_results_CONTAMINATED_2020.csv` |

The specification hash is verified at run time; the analysis refuses to execute
against a modified manifest.
