---
title: "Measuring geographic access with E2SFCA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Measuring geographic access with E2SFCA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette builds a four-tract, two-provider example small enough to check by
hand, and walks it through the same functions the national analysis uses. Every
number below is computed, not quoted, so if the engine changes the vignette
changes with it.

It uses no external data. The national surfaces depend on road-network
isochrones and ACS extracts that cannot ship in a package; the arithmetic does
not, and the arithmetic is what most people need to understand first.

```{r setup}
library(twostep)
```

## What the two steps are

Accessibility here is a **supply-to-demand ratio, smeared over a catchment**.

**Step 1** — for each provider *j*, add up the population that can reach it,
weighting nearer people more heavily, and divide that provider's capacity by the
total:

$$R_j = \frac{S_j}{\sum_i w_{ij} \, P_i}$$

**Step 2** — for each tract *i*, add up the ratios of every provider *i* can
reach, with the same weights:

$$A_i = \sum_j w_{ij} \, R_j$$

The weights $w_{ij}$ come from drive-time bands: someone 20 minutes away counts
more than someone 100 minutes away. That is the "enhanced" part — the original
two-step method treated everyone inside the catchment identically.

## Band weights, and why they are subtracted

Isochrone bands are **nested**: the 60-minute polygon contains the 30-minute
polygon. So the published weights are *cumulative*, and using them directly
would count the innermost population once per band it falls inside.

`e2sfca_incremental_weights()` performs the Abel summation that fixes this —
each ring gets the difference between its own weight and the next one out:

```{r weights}
W_cumulative <- c(`30` = 1.00, `60` = 0.68, `120` = 0.22, `180` = 0.09)
W_incremental <- e2sfca_incremental_weights(W_cumulative)
W_incremental
```

The outermost band keeps its cumulative weight, and the increments sum back to
the innermost cumulative weight — every ring counted exactly once:

```{r weights-identity}
sum(W_incremental)
identical(unname(sum(W_incremental)), unname(W_cumulative[["30"]]))
```

Get this wrong and accessibility is overstated everywhere, in a way that looks
entirely plausible. It is the single easiest place to introduce a believable
error, which is why the test suite pins it from several directions.

## A worked example

Four tracts, two providers. Tract populations and which band each tract falls
into for each provider:

```{r fixture}
# `pop_col` defaults to "female_pop", the denominator the study uses.
tract_pop <- data.frame(
  GEOID      = c("T1", "T2", "T3", "T4"),
  female_pop = c(1000, 2000, 1500, 500),
  stringsAsFactors = FALSE
)

# One row per (tract, provider, band). Nested bands: a tract in the 30-minute
# ring is also inside the 60-minute polygon, but appears once, at its tightest.
overlap <- data.frame(
  GEOID    = c("T1", "T2", "T3", "T4", "T2", "T3"),
  coord_id = c("P1", "P1", "P1", "P2", "P2", "P2"),
  band     = c(30L,  60L,  120L, 30L,  120L, 60L),
  # What fraction of the tract lies in that band. 1 = wholly inside.
  overlap_fraction = 1,
  stringsAsFactors = FALSE
)

supply <- data.frame(
  coord_id = c("P1", "P2"),
  supply   = c(4, 6),
  stringsAsFactors = FALSE
)
```

Step 1 by hand for provider P1: its weighted demand is
`1000 x 0.32 + 2000 x 0.46 + 1500 x 0.13`, and its ratio is `4` divided by that.

```{r step1-by-hand}
w <- W_incremental
demand_P1 <- 1000 * w[["30"]] + 2000 * w[["60"]] + 1500 * w[["120"]]
demand_P1
4 / demand_P1
```

Now the engine. Note it takes the **cumulative** weights and derives the
increments itself, so the correction above cannot be skipped by accident:

```{r run}
res <- compute_e2sfca(
  overlap   = overlap,
  tract_pop = tract_pop,
  supply    = supply,
  weights   = W_cumulative
)
res$access[, c("GEOID", "access", "access_math", "reached")]
```

## The conservation identity

The property worth internalising: **total supply is conserved**. Summing each
tract's accessibility, weighted by its population, returns total capacity —
accessibility redistributes supply across space, it does not create or destroy
it:

$$\sum_i P_i A_i = \sum_j S_j$$

This is the strongest single check on an accessibility surface. A result that
violates it is wrong no matter how plausible the map looks.

```{r conservation}
pop <- tract_pop$female_pop[match(res$access$GEOID, tract_pop$GEOID)]
c(redistributed = sum(pop * res$access$access_math),
  total_supply  = sum(supply$supply))
```

Those two agree exactly. Every tract with positive demand contributes; the identity holds over providers
reachable by someone. Tracts outside every catchment are a separate matter, and
the distinction is deliberate — see below.

## Unmeasured is not zero

A tract outside every modelled catchment has **no measurement**, which is not
the same as measured zero access. Reporting the two identically would let
missing data masquerade as a finding.

`compute_e2sfca()` therefore returns both:

- `access` is `NA` for tracts no provider reaches
- `access_math` is the zero-filled algebraic form, for sums and identities

Feed `access` to anything that averages and unreached tracts drop out; feed
`access_math` and they count as zero. Which you want depends on the question,
and the package refuses to choose for you.

## Missing data fails closed

The package will not silently repair an ambiguous join. If a tract has no
population row, or a population of `NA`, that is an error rather than a zero:

```{r fail-closed, error = TRUE}
tract_pop_gap <- tract_pop
tract_pop_gap$female_pop[tract_pop_gap$GEOID == "T3"] <- NA

compute_e2sfca(
  overlap   = overlap,
  tract_pop = tract_pop_gap,
  supply    = supply,
  weights   = W_cumulative
)
```

The reasoning is directional and worth stating plainly. Coercing unknown
population to zero removes people from the Step 1 denominator, which **inflates**
accessibility — the direction that makes a study look better than its data
supports. If an imputation is genuinely intended it can be requested by name:

```{r na-policy, eval = FALSE}
compute_e2sfca(
  overlap       = overlap,
  tract_pop     = tract_pop_gap,
  supply        = supply,
  weights       = W_cumulative,
  na_pop_policy = "zero"   # explicit, documented imputation
)
```

The same rule governs unmatched tracts, unmatched providers, and duplicate keys
throughout: **no silent loss, no silent multiplication, no silent zeroing.**
Duplicate keys are refused because a lookup would otherwise keep one row and
discard the rest, making the answer depend on row order. A duplicated supply row
was once found to *double* a tract's accessibility.

## Where to go next

| topic | see |
|---|---|
| the desjardins7 tract-level engine | `?dj7_tract_access` |
| the raster/grid path | `?build_e2sfca_grid_geometry`, `?compute_e2sfca_raster` |
| mass-conserving demand allocation | `?allocate_pop_areaweighted` |
| how figures and data were produced | `docs/FIGURE_PROVENANCE.md` |

The published parameterisation, the sensitivity specifications, and the national
results are in the manuscript, which reproduces from the frozen run.
